### 2020-05-22 paper.tex
3.1

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigplan,screen,10pt,anonymous,review]{acmart}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Customizing Software by Direct Manipulation of Tabular Data}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Geoffrey Litt}
\affiliation{%
  \institution{Massachusetts Institute of Technology}
  \city{Cambridge, MA}
  \country{USA}
}
\email{glitt@mit.edu}

\author{Daniel Jackson}
\affiliation{%
  \institution{Massachusetts Institute of Technology}
  \city{Cambridge, MA}
  \country{USA}
}
\email{dnj@csail.mit.edu}

\author{Tyler Millis}
\affiliation{%
  \institution{Massachusetts Institute of Technology}
  \city{Cambridge, MA}
  \country{USA}
}
\email{tmillis@mit.edu}

\author{Jessica Quaye}
\affiliation{%
  \institution{Massachusetts Institute of Technology}
  \city{Cambridge, MA}
  \country{USA}
}
\email{jquaye@mit.edu}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  In this paper we show how the behavior of a software application can
  be extended and adapted by direct manipulation. We introduce
  \emph{table-driven customization}, a new paradigm where end users
  directly manipulate a table view of the structured data inside an
  application, rather than writing imperative scripts (as in most
  customization tools). This simple model also accommodates a
  spreadsheet formula language and custom data editing widgets, which
  provide sufficient expressivity to implement many useful
  customizations.

  We illustrate the approach with Wildcard, a browser extension that
  implements table-driven customization in the context of existing web
  applications. Through concrete examples, we show that this paradigm
  can support useful customizations for real websites. We share
  reflections from our experiences using the Wildcard system, on both
  its strengths and limitations relative to other customization
  approaches. Finally, we explore how this paradigm might lead to new
  software architectures that encourage this form of customization.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
%% From HERE
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011066.10011069</concept_id>
<concept_desc>Software and its engineering~Integrated and visual development environments</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Integrated and visual development environments}
% To HERE

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{end-user programming, software customization, web browser extensions}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
%\begin{teaserfigure}
%  \includegraphics[width=\textwidth]{sampleteaser}
%  \caption{Seattle Mariners at Spring Training, 2010.}
%  \Description{Enjoying the baseball game from the third-base
%  seats. Ichiro Suzuki preparing to bat.}
%  \label{fig:teaser}
%\end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Many applications don't meet the precise needs of their users. End user
customization systems can help improve the situation, by empowering
non-programmers to modify their software in the way they would like.

Most end user customization systems offer a simplified version of
programming. Some scripting languages \citep{bolin2005, cook2007} have a
friendly syntax that resembles natural language. Other visual
customization tools eliminate text syntax entirely. Macro recorders
\citep{cook2007, chasins2018, anupam2000} remove some of the initial
programming burden by letting a user start with concrete demonstrations.
Despite their many differences, these approaches all share something in
common: an imperative programming model, with statement sequencing,
mutable variables and loops.

We have known for decades about an alternate approach: \emph{direct
manipulation} \citep{shneiderman1983}, where ``visibility of the object
of interest'' replaces ``complex command language syntax''. Direct
manipulation is the \emph{de facto} standard in GUIs today, but when it
comes to customizing those GUIs, it is rarely to be found. As a result,
switching from using software to customizing it poses a learning barrier
for users not familiar with programming, and requires an abrupt shift in
mental model even for those familiar with scripting.

In this work, we ask: what would it look like to build a direct
maniulation interface that lowers the threshold for starting to
customize software? We take inspiration from spreadsheets and visual
database query interfaces \citep{bakke2016, 2020a}, which have
successfully enabled millions of end users to compute with data through
direct manipulation.

In this paper we present a technique called \emph{table-driven
customization} inspired by those tools. An application's UI is augmented
with a table view, where the user can see and manipulate the
application's internal data. Changes in the table view result in
immediate corresponding changes to the original user interface, enabling
the user to customize an application with live feedback.

\begin{figure*}
\hypertarget{fig:overview}{%
\centering
\includegraphics[width=\textwidth]{media/overview.eps}
\caption{An overview of table-driven customization}\label{fig:overview}
}
\end{figure*}

We have developed a browser extension called Wildcard which uses web
scraping techniques to implement table-driven customization for existing
Web applications. In Section~\ref{sec:example}, we illustrate the end
user experience of table-driven customization through an example
scenario using Wildcard.

In Section~\ref{sec:architecture}, we explain the system architecture of
table-driven customization. We focus on the \emph{table adapter}
abstraction, which allows many different types of underlying data to be
bidirectionally mapped to a table. We describe several types of table
adapters we've built in Wildcard, and also describe future adapters
supported by the general paradigm.

We have successfully used Wildcard to build customizations for 11
different websites. In Section~\ref{sec:evaluation}, we present
reflections from this process, outlining the kinds of customizations we
were able to build, limitations we encountered, and the challenges of
integrating scraping logic with websites.

In Section~\ref{sec:themes}, we discuss some key themes from our work:

\begin{itemize}
\tightlist
\item
  \emph{Customization by direct manipulation}: We agree with MacLean et
  al \citep{maclean1990} that customization should follow a ``gentle
  slope'', where users can gradually invest in more complex techniques
  to achieve deeper customizations. We believe an important point on
  this slope is the ability to customize an application by directly
  seeing and changing its internal structured data, rather than by
  writing imperative scripts.
\item
  \emph{Wrapping applications for customization}: Typically, tools that
  don't rely on official extension APIs resort to offering low-level
  APIs for customization. Instead, we propose a community-maintained
  library of semantic wrappers around existing applications, enabling
  end users to work with domain objects rather than low-level
  representations.
\end{itemize}

Table-driven customization relates to existing work in many areas. In
particular, our goals overlap with many software customization tools,
and our methods overlap with direct manipulation interfaces for working
with structured data, including visual database query systems and
spreadsheets. We explore these connections and more in
Section~\ref{sec:related-work}.

\hypertarget{sec:example}{%
\section{Example Scenario}\label{sec:example}}

To illustrate the end user experience of table-driven customization, we
consider an example scenario of customizing Hacker News, a tech news
aggregator. Figure~\ref{fig:hacker-news} shows accompanying screenshots.

\begin{figure*}
\hypertarget{fig:hacker-news}{%
\centering
\includegraphics[width=\textwidth]{media/hacker-news.png}
\caption{Customizing Hacker News by interacting with a table view}\label{fig:hacker-news}
}
\end{figure*}

\textbf{Opening the table}: When the user opens Hacker News in a browser
equipped with the Wildcard extension, they see a table at the bottom of
the page. It contains a row for each link on the homepage, listing
information like the title, URL, submitter username, number of points,
and number of comments (Figure~\ref{fig:hacker-news}, Note A). The end
user didn't need to do any work to create this table, because a
programmer previously created an adapter to extract data from this
particular website, and contributed it to a shared library of adapters
integrated into Wildcard.

\textbf{Sorting by points}: First, the user decides to change the
ranking of links on the homepage. Hacker News itself uses a ranking
algorithm in which the position of an article depends not only on its
point count (a measure of popularity), but also on how long it has been
on the site. If the user hasn't been checking the site frequently, it's
easy to miss a popular article that has fallen lower on the list.
Sorting the page just by points would achieve a more stable ranking.

To achieve this ordering, the user simply clicks on the ``points''
column header in the table. This sorts the table view by points, and the
website UI also becomes sorted in the same order
(Figure~\ref{fig:hacker-news}, Note B)---Wildcard has manipulated the
website's DOM to synchronize it with the sort order of the table. This
sort order is also persisted in the browser and reapplied automatically
the next time the user loads the page, so they can always browse the
latest set of links using this sort order.

\textbf{Adding estimated read times}: Next, the user decides to attempt
a more substantial customization: to add estimated read times to
articles on Hacker News.

The table contains additional columns to the right of the original
attributes for each row, where the user can enter spreadsheet-style
formulas to compute derived values. We provide generic names for these
columns by default as spreadsheets do, so that users don't need to
choose column names themselves. The user enters a formula into the
\texttt{user1} column: (Figure~\ref{fig:hacker-news}, Note C):

\begin{verbatim}
=ReadTimeInSeconds(link)
\end{verbatim}

This calls a built-in function that takes a URL as an argument and uses
a public web API to compute an estimated read time for the link's
contents. The argument \texttt{link} refers to a column name in the
table; for each row, the formula is evaluated with the link for that
particular row. The result of evaluating this formula is a numerical
value in seconds for each article.

The user clicks the \texttt{user1} column header to sort the articles on
the page in descending order of estimated read time, helping to
prioritize reading deeper content. It would be useful to show these read
times in the page too, but before doing that, the user wants to make the
time estimates more readable. They enter another formula in the next
column, \texttt{user2}:

\begin{verbatim}
=Concat(Round(user1/60), "min read")
\end{verbatim}

This formula converts seconds to minutes by dividing by 60 and rounding
to the nearest integer, and then concatenates a label to the number,
producing strings like ``21 min read''.

Finally, the user clicks a menu option in the table header to display
the contents of this new column in the original page
(Figure~\ref{fig:hacker-news}, Note D). Each article on the page now
shows an annotation with the estimated read time in minutes. (The format
of how annotations appear for each item was determined by the programmer
who created the Wildcard adapter for Hacker News.)

\textbf{Adding notes to links}: The user can also manually add notes to
the table, by simply entering values into the table without using
formulas. In this case, the user jots down a few notes in another column
about articles they might want to read, and the notes appear in the page
next to the read times (Figure~\ref{fig:hacker-news}, Note D).

\textbf{Filtering out visited links}: Another way to use formulas to
customize Hacker News is to filter out articles the user has already
read. (We omit this example from the figure for brevity.) The user can
call a built-in function that returns a boolean depending on whether a
URL is in the browser's history:

\begin{verbatim}
=Visited(link)
\end{verbatim}

They can then filter the table to only contain rows where this formula
column contains \texttt{false}; visited rows are hidden both from the
table view and the original page. This is an example of a customization
that the original website could not implement, since websites don't have
access to the browser history for privacy reasons. But by using
Wildcard, the user can implement the customization locally, without
needing to expose their entire browser history to Hacker News.

This scenario has shown a few examples of how table-driven
customizations can help a user improve their experience of a website.
Section~\ref{sec:evaluation} explains many other use cases, but first we
explain how the system works internally.

\hypertarget{sec:architecture}{%
\section{System architecture}\label{sec:architecture}}

\begin{figure}
\hypertarget{fig:table-adapter}{%
\centering
\includegraphics[width=\columnwidth]{media/table-adapter.eps}
\caption{The table adapter architecture}\label{fig:table-adapter}
}
\end{figure}

Figure~\ref{fig:table-adapter} summarizes the overall architecture of
table-driven customization, using a simplified version of the Airbnb
example above. (\emph{todo: change this figure to hacker news}) In this
example, the name of each listing is scraped from the web page DOM, the
latitude and longitude of each listing is scraped from AJAX responses,
and user annotations are loaded from the brower's local storage.

First, the three different data sources are each bidirectionally mapped
to a table interface by a \textbf{table adapter}. The table adapter
defines how to map a particular type of data to a table, and what
effects edits should have on the original data source. In some cases,
the mapping logic is straightforward: the local storage adapter stores a
table of data, so the mapping to the table abstraction is trivial. In
other cases, the mapping is more involved: the DOM scraping adapter
implements web scraping logic to produce a table of data from the web
page, and turns edits to the table into DOM manipulations like
reordering rows of data on the page.

The three separate tables are then combined into a single table for the
end user to view and edit. The \textbf{query engine} is responsible for
creating this combined view, and routing the user's edits back to the
individual table adapters. In this example, the query engine has joined
the three tables together by a shared ID column, and sorted the result
by the name column.

We now examine each component of the system in more detail.

\hypertarget{table-adapters}{%
\subsection{Table adapters}\label{table-adapters}}

A key idea in table-driven customization is that a wide variety of data
sources can be mapped to the generic abstraction of an editable table.
In a relational database, the table matches the underlying storage
format, but in table-driven customization, the table is merely an
\emph{interface layer}. The data shown in the table is a projection of
some underlying state, and edits to the table can have complex effects
on the underlying state.

Externally, a table adapter must satisfy an abstract interface. The
first two parts of the interface resemble the interface exposed by a
table in a relational database:

\textbf{Returning a table}: A table adapter exposes a table of data: an
ordered list of records. Each record carries a unique identifier and
associates named attributes with values. Tables have a typed schema, so
the same attributes are shared across all records. A table adapter can
update the contents of a table at any time, in response to changes in
the underlying state (e.g., a DOM scraping adapter can update the table
when the page body changes). When data changes, the query view is
reactively updated in response.

\textbf{Making edits}: The query engine can issue a request to a table
adapter to make an edit to a record. The meaning of making an edit can
vary depending on the adapter: in the local storage adapter, an
annotation may be persisted into local storage; in the DOM scraping
adapter, an edit may represent filling in a form field. An adapter can
also mark values as read-only if it wouldn't be meaningful to edit them;
for example, the DOM scraping adapter typically marks page content as
read-only, except for editable form fields.

The query engine also sends additional information about the combined
query view to each table adapter. These functions are currently used to
provide the DOM scraping adapter with sufficient information to
manipulate the original application UI as the user manipulates the table
view.

\textbf{Sorting/filtering}: When the user sorts or filters the query
view, an ordered list of visible IDs is sent to each table adapter. The
DOM scraping adapter uses this information to change the list of rows
shown in the web page.

\textbf{Information from other tables}: The query engine sends each
table adapter the entire combined table of data being rendered to the
user. The DOM scraping adapter uses this to add annotations to the page
by looking for additional data columns joined to scraped rows, and
rendering them in the page. (\emph{todo: explain joined})

\textbf{Currently selected record}: The query engine notifies each table
adapter about which record is currently selected by the user in the
table UI. The DOM scraping adapter uses this information to highlight
the row in the page that corresponds to the selected record in the
table. (\emph{todo: clarify})

\hypertarget{types-of-table-adapters}{%
\subsubsection{Types of table adapters}\label{types-of-table-adapters}}

Here we describe in more detail the table adapters that we have
implemented in Wildcard to power the customizations shown in
Section~\ref{sec:example}.

\textbf{DOM scraping adapters} are the essential component that enables
Wildcard to interface with an existing website UI. In addition to the
standard web scraping problem of extracting a table of data from the
DOM, a scraping adapter must also manipulate the DOM to reorder rows,
edit form entries, and inject annotations as the table is edited.

Because each website has unique content, we rely on programmers to
create a DOM scraping adapter for each individual website to make it
available for customization in Wildcard. To make this approach viable,
we have built a library of functions that make it easy to implement a
scraping adapter, requiring the programmer only to implement the minimal
site-specific parts. The programmer writes a single Javascript function
which can use standard scraping techniques like CSS selectors and DOM
APIs to extract relevant elements from the page. The library functions
then wrap this function to implement the rest of the needed
functionality; for example, when the table is sorted, we find the DOM
elements corresponding to the rows in the table, remove them from the
page, and then reinsert them in the new order.

An \textbf{AJAX scraping adapter} intercepts AJAX requests made by a web
page, and extracts information from those requests. When available, this
tends to be a helpful technique because the data is already in a
structured form, and often includes information not shown in the UI. As
with DOM scraping adapters, we have made it easy for programmers to
create site-specific AJAX scraping adapters. A programmer writes a
function that specifies how to extract data from an AJAX request, and
the framework handles the details of intercepting requests and calling
the programmer-defined function.\footnote{So far we have only
  implemented AJAX scraping in the Firefox version of Wildcard, since
  Firefox has convenient APIs for intercepting requests. It appears
  possible to implement in Chrome as well, but we have not finished our
  implementation.}

The \textbf{local storage adapter} simply stores a table of data in the
browser. As shown in the example use cases, this can be useful for
persisting annotations in a private way, without uploading them to a web
service.

\hypertarget{future-adapters}{%
\subsubsection{Future Adapters}\label{future-adapters}}

We have designed the table adapter API to be general enough to support
other types of useful adapters in the future. Here are two such
possibilities:

\textbf{Integrated website adapters}: A key benefit of the table adapter
abstraction is that Wildcard is not coupled to web scraping as the only
means for integrating with existing sites, but can also accommodate
first party developers adding support directly into their own websites.
An ``integrated website adapter'' installed by the developer could
directly accesses the internal state of the application, providing the
same functionality as a DOM scraping adapter but in a more robust way.

With the advent of rich frontend web frameworks, structured application
state is now often available in the web client. We suspect it is
possible to create plugins for frontend frameworks that expose this
state to Wildcard with only minimal effort from the application
developers. To test this hypothesis, we have created an early prototype
of a Wildcard adapter for the Redux state management library, and used
it with the TodoMVC todo list application. To install the adapter, the
programmer only needs to make a few small changes to their app:

\emph{Exposing a table}: In Redux, app developers already represent the
state of a user interface as a single object. To expose a table of data,
the developer simply writes a function that maps the state object to a
table. \emph{Editing data values}: The developer defines how edits to
records in the table should map to edits in the state object.
\emph{Sorting/filtering the UI}: The developer inserts an additional
data processing step before the state object is rendered in the UI, so
that Wildcard can apply sorting and filtering. \emph{Annotations}: The
developer adds additional React components managed by Wildcard into
their UI tree, which render additional data as annotations.

In certain types of software, customizability is often touted as a key
selling point. An integrated website adapter would provide a way for
developers to integrate with an existing ecosystem of formulas and
customization tools, without needing to build all that functionality
from scratch.

\textbf{Shared storage adapter}: It would be useful to share user
annotations among users and across devices---for example,
collaboratively taking notes with friends on options for places to stay
on Airbnb. The existing Local Storage Adapter could be extended to share
live synchronized data with other users. This could be achieved through
a centralized web server, or through P2P connections that might provide
stronger privacy guarantees.

\hypertarget{query-engine}{%
\subsection{Query engine}\label{query-engine}}

The query engine is responsible for coordinating across multiple table
adapters. It joins data across multiple tables and creates a single
result table which is shown to the user through the editor. It also
handles all user interactions and routes appropriate messages to each
table adapter.

Queries are processed in three steps. First, the query invokes a primary
DOM scraping table adapter that associates records in the result with
elements in the application's user interface. At minimum, the primary
table adapter needs to return record IDs and have the ability to
manipulate the application's UI. It can also optionally return data
about each record. Next, additional tables (AJAX data, local storage
data) are left-joined by ID. Finally, the result table can be sorted and
filtered by any column.

This query model can be viewed as a tiny (and rather constrained) subset
of the SQL query model. Despite its simplicity, this simple model has
proven sufficient for meeting the needs of customization in practice,
and minimizes the complexity of supporting more general and arbitrary
queries. But because it fits into the general paradigm of relational
queries, it could theoretically be extended to support a wider range of
queries.

The query engine is also responsible for executing formulas. We have
built a small formula language resembling the ones used in visual query
tools like SIEUFERD. As in those tools, and unlike in spreadsheets,
formulas automatically apply across an entire column of data, and
reference other column names instead of values on specific rows. For
example, the user would write \texttt{=price\ *\ 2} to create a column
with a value derived from a price attribute. This is more convenient
than needing to copy-paste a formula across an entire column as in
spreadsheets, and none of our customization use cases have uncovered a
need for writing different formulas on different rows of the table.

\hypertarget{table-editor}{%
\subsection{Table editor}\label{table-editor}}

In Wildcard we provide a basic table editor as the user interface on top
of the query engine. It is built with the Handsontable Javascript
library, which provides UI elements for viewing and editing a table, as
well as basic query operations like sorting and filtering.

\emph{todo: other table instruments}

\begin{table*}[]
\begin{tabular}{llrl}
\hline
\textbf{Website} & \textbf{Description} & \textbf{LOC} & \textbf{Example customizations}                                                              \\ \hline
Airbnb           & Travel               & 73                                       & Add Walkability Scores to listings. Sort listings by price.                           \\
Amazon           & Online shopping      & 99                                       & Sort used book sellers by total price, including delivery fees                                \\
Blogger          & Blogging             & 36                                       & Use alternate text editor to edit blog posts                                                 \\
Expedia          & Travel               & 41                                       & Use alternate datepicker to enter travel dates                                               \\
Flux             & Data portal          & 67                                       & Use Wildcard as a faster table editor for editing lab results                                \\
Github           & Code repository      & 62                                       & Sort a user's code repositories by stars to find popular work                                \\
Hacker News      & News                 & 69                                       & Add read times to links. Filter out links that have been read. \\
Instacart        & Grocery delivery     & 48                                       & Sort groceries by price and category. Take notes on items.                                   \\
Uber Eats        & Food delivery        & 117                                      & Sort/filter restaurants by estimated delivery ETA and price.                                 \\
Weather.com  & Weather              & 51                                       & Sort/filter hourly weather to find nice times of day.                                        \\
Youtube          & Videos               & 80                                       & Sort/filter videos by length, to find short videos to watch.                                 \\ \hline
\end{tabular}
\caption{A list of table-driven customizations that we have implemented using Wildcard.}
\label{tab:websites}
\end{table*}

\hypertarget{sec:evaluation}{%
\section{Evaluation}\label{sec:evaluation}}

To evaluate table-driven customization in practice, we built the
Wildcard browser extension, which implements table-driven customization
in the context of existing websites. It is implemented in Typescript,
and works across three major browsers: Chrome, Firefox, and Edge.

We have used Wildcard to customize 11 websites, including transactional
sites like Amazon and Uber Eats, and media consumption sites like Hacker
News and Youtube. Table 1 summarizes these customizations, showing the
number of lines of code in the adapter configuration for each site.

So far, most usage of Wildcard has come from members of the project
team. Here we offer our reflections on using the system, focused on two
key questions:

\begin{itemize}
\tightlist
\item
  How broad is the range of possible customizations in this paradigm?
\item
  How feasible is it to build DOM scraping adapters for real websites?
\end{itemize}

\hypertarget{range-of-customizations}{%
\subsection{Range of customizations}\label{range-of-customizations}}

We have found that table-driven customization can serve a broad range of
useful purposes. Here we expand on some archetypal examples that
illuminate various aspects of using the system in practice.

\hypertarget{sortingfiltering}{%
\subsubsection{Sorting/filtering}\label{sortingfiltering}}

It might seem that most websites already have adequate sorting and
filtering functionality, but we have found it surprisingly helpful to
add new sorting/filtering functionality to websites using Wildcard.

Sometimes, websites have opaque algorithms which they use to rank
results, which presumably maximize profit but don't offer much control
to the user. For example, Airbnb doesn't allow users to sort listings by
price, and Youtube doesn't allow users to sort the homepage by watch
time. Wildcard enables users to take back some control over this
process.

In other cases we have found that a lack of sorting options seems less
intentional, and simply represents a missing feature. For example, the
Instacart grocery delivery service doesn't allow users to sort their
grocery cart by price or category. Our guess is that they simply haven't
gotten around to implementing the feature yet.

In the current implementation of Wildcard, users can only sort and
filter entries that are shown on the current page, which means that
users are not entirely liberated from the suggestions of the opaque
algorithm. We could work around this restriction in the future by
fetching content across multiple pages, but we've also found that
sorting/filtering a single page of a paginated list is often an
acceptable outcome (and sometimes even a preferable one). For example,
it's more useful to sort (by length, say) 30 recommended Youtube videos
than to try to sort all videos on Youtube, or to sort (by read time,
say) the articles on the Hacker News front page than to sort all
non-archived articles.

\hypertarget{annotating}{%
\subsubsection{Annotating}\label{annotating}}

Many other web annotation systems focus on annotating text or arbitrary
webpage content, but Wildcard limits annotations to structured objects
extracted by an adapter, resulting in quite a different set of use
cases.

Annotating with Wildcard has proven most useful when taking notes on a
list of possible options (e.g., evaluating possible Airbnb locations to
rent). We have also used it with Instacart's online grocery cart, for
jotting down notes like ``should we get more milk?''

\hypertarget{formulas}{%
\subsubsection{Formulas}\label{formulas}}

Formulas are the most powerful part of the Wildcard system. So far, our
language supports only a small number of predefined functions. Adding
more will allow a very broad range of useful computations, as shown by
the success of spreadsheets.

Formulas are especially useful for fetching data from Web APIs. We've
used them to augment Airbnb listings with walkability scores, and to
augment Hacker News articles with estimated read times. One challenge of
the current language design is that supporting a new web API requires
writing conventional code to add a new function to the language, because
web APIs typically return complex JSON data structures that can't be
easily displayed in a single table cell. Adding functions for handling
JSON data to the formula language might make it possible to access new
APIs without requiring an entire new function in the formula system.

We have also found instances where simple data manipulation is useful:
adding together different subtotals into a total price, or manipulating
the results of a formula expression with arithmetic or concatenating
string labels, as shown in the example in Section~\ref{sec:example}.
(\emph{todo: this makes sense iff we switch the examples sectino to HN})

\hypertarget{cell-editors}{%
\subsubsection{Cell editors}\label{cell-editors}}

(\emph{note: written with the assumption that Expedia gets cut from the
Examples section})

Cell editors are UI widgets that expose a custom editing UI for a single
cell of the table view. A programmer building a cell editor need only
integrate it with the table viewer; propagating values into the website
UI is handled by the site-specific DOM adapter. As a result, creating a
new cell editor can be as easy as just writing some glue code between
the table editor and an existing UI widget library.

An important benefit of cell editors is that they allow users to
incorporate their personal information within the web UI but without
uploading their personal data to the web server. To explore this idea,
we created a cell editor based on the FullCalendar Javascript plugin,
which can load data from a Google Calendar. This makes it convenient to
enter dates into a website based on the user's personal calendar
information.

Another benefit of cell editors is that a user can choose their
preferred widget for editing some type of information. We built a cell
editor based on the CKEditor rich text editor, and demonstrated its use
by integrating it with Google's Blogger website and using it to edit
blog posts, in place of Blogger's built-in editor.

\hypertarget{limitations}{%
\subsubsection{Limitations}\label{limitations}}

There are many useful customizations that are not currently possible in
the table-driven customization paradigm. Wildcard can only make
customizations that use the available data exposed in the table. UI
modifications are limited in scope; deleting arbitrary buttons isn't
possible, for example. At one point, we wanted to build an automation to
repeatedly load a grocery delivery website to check for open delivery
slots, but it wasn't clear how to achieve this. Some of these
limitations are specific to the current implementation of the Wildcard
extension, but some are more fundamental to the entire idea of
table-driven customization.

We consider this an acceptable outcome; our goal is to make many useful
customizations available to end users with a low threshold, and not to
span all possible customizations. Section 5.1 discusses this point
further.

One benefit of Wildcard is predictability: once we built a site adapter
for a website, it was generally obvious to us what types of
customizations were and weren't possible, and the UI guided us towards
building customizations that matched the system's model. If data is not
available in the table, for example, then it's abundantly clear that
customizations relying on that data will not be expressible.

\hypertarget{viability-of-dom-scraping}{%
\subsection{Viability of DOM scraping}\label{viability-of-dom-scraping}}

Our second evaluation area relates less to the conceptual approach of
table-driven customization, and more to the specific implementation of
customizing existing web applications. In order for third-party
customization through Wildcard to succeed, it is important that creating
usable adapters for existing websites takes minimal effort.

Most of our DOM scraping adapters were created by members of our team.
However, an external developer unaffiliated with the project contributed
one adapter, designed to sort the Github page listing a user's
repositories. They described the experience as ``very straightforward.''

The adapters for our test sites ranged from 36 to 117 lines of code,
averaging 68 lines; Table 1 shows the full lines of code for each
adapter.

Some of the challenges of writing a DOM scraping adapter are the same
ones as with writing normal web scraping code, but the more interactive
nature of Wildcard introduces additional challenges. rve changes, but it
may prove challenging to observe changes on some sites only through the
DOM. One challenge is triggering updates to the spreadsheet data in
response to UI changes that happen after initial page load. Site
adapters are responsible for recognizing these changes by observing the
DOM. So far, we have been able to use event listeners and the
MutationObserver API to successfully obse

Another challenge is persisting updates to the DOM---some websites use
virtual DOM frameworks that can occasionally overwrite changes made by
Wildcard.

So far, we've managed to work around these issues for all the websites
we've tried, but we don't claim that any website can be customized
through DOM scraping. As web frontend code gets increasingly complex
(and starts to move beyond the DOM to other technologies like Shadow DOM
or even WebGL for rendering), it may become increasingly difficult to
customize websites from the outside without first-party support.

\hypertarget{sec:themes}{%
\section{Key themes}\label{sec:themes}}

\hypertarget{sec:dm}{%
\subsection{Customization by direct manipulation}\label{sec:dm}}

Hutchins, Hollan and Norman \citep{hutchins1985} define a direct
manipulation interface as one that uses a model-world metaphor rather
than a conversation metaphor. Instead of presenting an ``assumed'' but
not directly visible world that the user converses with, ``the world is
explicitly represented'' and the user can ``{[}act{]} upon the objects
of the task domain themselves.''

Although most GUIs today employ direct manipulation, software
customization tools typically use an imperative programming model, which
implements the conversational metaphor rather than direct manipulation.
Here, for example, is how a user retrieves a list of of calendar names
from the Calendar application in Applescript \citep{cook2007}, the
scripting language for customizing Mac OS applications:

\begin{verbatim}
tell application "Calendar"
  name of calendars
end tell
\end{verbatim}

Some customization environments like Mac Automator and Zapier forego
textual syntax and let the user connect programs and construct
automations by dragging and dropping icons representing commands. These
environments still do not constitute direct manipulation, though: the
objects being manipulated are in the domain of programming, not in the
domain of the task at hand. Imperative programming is a reasonable
choice as the model for building customizations. Turing-complete
programming provides a high ceiling for possible customizations, and a
sequence of commands is a natural fit for automations that simulate a
series of steps taken by the user.

Imperative programming is a reasonable choice as the model for building
customizations. Turing-complete programming provides a high ceiling for
possible customizations, and a sequence of commands is a natural fit for
automations which simulate a series of steps taken by the user.

There is, however, a serious drawback to this approach. MacLean et al
\citep{maclean1990} describe an ideal for user-tailorable systems: a
``gentle slope'' from using to customizing, where small incremental
increases in skill lead to corresponding increments of customization
power. Requiring users wanting to customize their applications to learn
programming creates an abrupt ``cliff,'' exacting a significant
investment in learning even to implement the simplest customizations.
Another goal of MacLean et al is to make it ``as easy to change the
environment as it is to use it''---at least for some subset of changes.
But in scripting languages, even user-friendly ones like Applescript or
Chickenfoot, the experience of customization does not remotely resemble
the experience of use, so these systems can't meet this goal.

With table-driven customization we aim to provide a gentler slope, by
using direct manipulation for software customization. The data shown in
the table view is the domain data from the original application. The
user makes changes to the data by selecting areas of interest in the
table, e.g.~sorting/filtering by clicking the relevant column header, or
adding annotation by clicking on the relevant row. These interactions
are common in GUI applications, and Wildcard therefore meets MacLean et
al's goal: some one-click customizations are as easy as using the
original application.

One aspect of directness that we have chosen not to maintain in Wildcard
is enabling customization in the context of the original user interface,
as explored by tools like Scotty \citep{eagan2011}. We have found that
augmenting the original UI with a distinct, additional representation
provides a consistent experience across all applications, and clearly
shows the user what structured data is available to work with.

Ainsworth et al provide a helpful taxonomy of the value of multiple
representations \citep{ainsworth1999}. In their terms, Wildcard plays a
complementary role by supporting a different set of tasks from the
original application, while displaying shared information. Wildcard may
also help construct deeper understanding by subtraction. By stripping
away details and only showing the essential data in an interface,
Wildcard helps users think of new ways of using that data, outside the
specific constraints of the original application.

It is important to note that there are many specific customizations that
can be achieved in systems like Chickenfoot \citep{bolin2005} that
cannot be reproduced in Wildcard. We consider this an acceptable
tradeoff in exchange for a gentler slope in customization, and we show
in Section~\ref{sec:example} and Section~\ref{sec:evaluation} that our
model can still implement many useful customizations in practice. Also,
there is sometimes a way to reframe an imperative script in terms of our
direct manipulation model; for example, a script that iterates through
rows in a page adding some additional information can be reproduced
using a formula in Wildcard.

\hypertarget{semantic-wrappers}{%
\subsection{Semantic wrappers}\label{semantic-wrappers}}

Software customization tools typically fall into one of two categories:
ad hoc, or semantic.

\emph{Ad hoc customization tools} enable customization without using
official extension APIs, enabling a broader range of customizations on
top of more applications. For example, web browser extensions have
demonstrated the utility of customizing websites through manipulating
the DOM, without needing explicit extension APIs to be built in.
However, ad hoc customization comes with a corresponding cost: these
tools can typically only operate at a low level of abstraction,
e.g.~manipulating user interface elements. This makes it harder for end
users to write scripts, and makes the resulting scripts more brittle.

\emph{Anticipated customization tools}, in contrast, use explicit
extension APIs provided by the application developer. Examples of this
include accessing a backend web API, or writing a customization in
Applescript that uses an application-specific API. The main benefit is
that this allows the extension author to work with meaningful concepts
in the application domain---``create a new calendar event'' rather than
``click the button that says new event.''---which makes customizations
easier to build and more robust. However, this style limits the range of
extensions that can be built to those that the official plugin API
supports.

\emph{todo: copy over more content from gdoc}

With Wildcard, we use a hybrid approach that aims to provide the best of
both worlds. Programmers implement an API wrapper that is internally
implemented as an ad hoc customization, but externally provides a
high-level interface to the application, abstracting away the details of
the user interface. These wrappers are added to a shared repository,
available to all users of the system. When an end user is using a site
that already has an adapter, they get a semantic customization
experience that avoids low-level details.

One way to view this approach is as introducing a new abstraction
barrier into third-party extension. Typically, a third party
customization script combines two responsibilities: 1) mapping the
low-level details of a user interface to semantic constructs (e.g.,
using CSS selectors to find certain page elements), and 2) handling the
actual logic of the specific customization. Even though the mapping
logic is often more generic than the specific customization, the
intertwining of these two responsibilities in a single script makes it
very difficult to share the mapping logic across scripts.

With Wildcard we propose a decoupling of these two layers: a repository
of shared wrappers maintained by programmers, and a separate repository
of specific customizations built on top of these wrappers. This general
architecture has been successfully demonstrated by projects like
Gmail.js, an open source project that creates a convenient API for
browser extensions to interface with the Gmail web email client.

(\emph{todo: mention automated wrapper induction here or somewhere
else?})

\hypertarget{sec:related-work}{%
\section{Related Work}\label{sec:related-work}}

This paper extends work reported in a workshop paper by Litt and Jackson
\citep{litt2020} which presented an early prototype version of Wildcard.
We have substantially extended their work in this paper by creating the
table adapter abstraction and reimplementing the system around that
abstraction, evaluating the system more fully on many more websites, and
by characterizing the design of the system in more detail.

Table-driven customization relates to two broad areas of related work.
Our problem statement is related to software customization tools, and
our solution approach is related to spreadsheets and other direct
manipulation interfaces.

\hypertarget{customization-tools}{%
\subsection{Customization tools}\label{customization-tools}}

Table-driven customization is most closely related to other tools that
aim to empower end users to customize software without traditional
coding.

This lineage goes back at least to the Buttons system by MacLean et al
\citep{maclean1990}, where Xerox Lisp users could share buttons that
performed various ``tailoring'' actions on the system. The authors
proposed the ``gentle slope'' idea which has greatly influenced our
approach to table-driven customization (as discussed in
Section~\ref{sec:dm}). The authors also point out the importance of a
``tailoring culture'' where people with different skillsets collaborate
to produce useful customizations; in their system, Lisp programmers
create buttons that others can use, modify, and rearrange. This division
of labor corresponds to our idea of semantic wrappers, where end user
customization is supported by programmer-created building blocks.

Some recent web customization tools aim to enable end users to modify
web interfaces without programming.

Sifter \citep{huynh2006} enables end users to sort and filter lists of
data obtained by web scraping, much like Wildcard's sorting features.
The main difference between the systems is that table-driven
customization has many other use cases besides sorting and filtering.
Also, Sifter involves end users in a semi-automated data extraction
process, rather than having programmers create wrappers. This provides
coverage of more websites, but at the expense of complicating the end
user experience. We might explore such techniques in Wildcard in the
future, but we think that it's valuable for end users to have a
customization experience decoupled from the challenge of web scraping
the underlying data. Sifter implements scraping across multiple pages, a
valuable feature for sorting and filtering that isn't present in
Wildcard.

Thresher \citep{hogue2005} helps end users create wrappers that map
website content to Semantic Web schemas, making customizations available
using the schema information. Wildcard shares the general idea of
wrappers with Thresher, but maps to a single generic data type, rather
than more specific schemas, increasing the range of websites and data
that Wildcard is able to support.

There are many software customization tools that offer simplified forms
of programming for end users. Chickenfoot \citep{bolin2005} and
Coscripter \citep{leshed2008} offer user friendly syntax for writing web
automation scripts; Applescript \citep{cook2007} has a similar goal for
desktop customization. There are visual programming environments for
customization that don't involve writing any text:
\href{https://support.apple.com/guide/automator/welcome/mac}{Automator}
for Mac and
\href{https://apps.apple.com/us/app/shortcuts/id915249334}{Shortcuts}
for iOS are modern options for customizing Apple products, and
\href{https://zapier.com/}{Zapier} enables users to connect different
web applications together visually. As mentioned previously, these tools
all require writing imperative programs in some form, in contrast to the
more declarative and direct approach of table-driven customization.

\hypertarget{direct-manipulation-programming-interfaces.}{%
\subsection{Direct manipulation programming
interfaces.}\label{direct-manipulation-programming-interfaces.}}

Another relevant area involves direct manipulation interfaces for
interacting with structured data. We take inspiration from these tools
in our work, but apply them in a different domain: customizing existing
software applications, rather than interacting with databases or
constructing software from scratch.

The most closely related work is in systems that offer spreadsheet-like
querying of relational data, as proposed by Liu and Jagadish
\citep{liu2009}. SIEUFERD by Bakke and Karger \citep{bakke2016} is an
example; their paper presents a survey of other similar tools. Our work
is particularly influenced by the authors' observation that direct
manipulation requires that the user manipulate the results of a database
query rather than the query itself, and that the user must see
intermediate results at every step of constructing the query. SIEUFERD's
interface supports a far wider range of queries than Wildcard, but the
basic ideas of the user interface are similar.

\href{https://airtable.com/}{Airtable} is another example of a modern
commercial product that offers spreadsheet-like interaction with a
relational database.

Our work is also inspired by the many projects that have explored using
spreadsheets as a foundation for building software applications,
including Object Spreadsheets \citep{mccutchen2016}, Quilt
\citep{benson2014}, Gneiss \citep{chang2014}, Marmite \citep{wong2007},
\href{https://www.glideapps.com/}{Glide}. These projects share the idea
that a spreadsheet is a convenient interface for editing the data
underlying a GUI application. We share that idea, but apply it to
software customization, rather than building software from scratch.

Another system in this space is ScrAPIr, by Alrashed et
al.~\citep{alrashed2020}, which enables end users to access backend web
APIs without programming. ScrAPIr shares our high level goal of end user
empowerment, as well as the idea of wrappers, by creating a shared
library of wrappers around existing web APIs. Unlike Wildcard, however,
ScrAPIr targets explicit APIs exposed by developers. It also focused on
backend services and doesn't aim to extend the frontend interfaces of
web applications.

\hypertarget{sec:conclusion}{%
\section{Conclusion and Future Work}\label{sec:conclusion}}

In this paper, we have presented table-driven customization, a new
paradigm for customizing software by direct manipulation of the
underlying structured data. We have demonstrated the paradigm using a
browser extension, and have used it to create useful customizations for
a variety of websites.

Our primary goal for future work is to evaluate the system with a
broader group of users. What barriers do end users face when using the
system? What types of customizations do they choose to create? What
formulas prove most helpful?

Another area for future work is exploring whether it is possible to
express a broader range of customizations by extending our table-editing
paradigm. Are there ways to offer an increase in power and functional
complexity, while retaining a programming model that is simpler for end
users than conventional coding?

As computing plays an ever greater role in our lives, it is increasingly
important that end users have agency over the behavior of their
software, rather than having every detail be dictated by companies whose
incentives are not always aligned with the user's interests. We hope
that table-driven customization can serve as one point, and maybe a
guidepost, on our path from normal use to deep modification, in support
of a more adaptable experience for all computer users.

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{wildcard-onward-bibtex.bib}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.